---
title: 'DATA 252 / DATA 551: Homework 3 Solution'
author: ''
date: ' '
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
fontsize: 12pt
---

<!-- do not alter this first chunk below  -->
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\vspace{0.1in}
\noindent 1. In \textit{Introduction to Probability}, skim through pages 183 to 187; don't worry about the technical details, but make sure that you understand figures 5.1 and 5.2.  Read the section on \textbf{Possin Distribution}, pages 187 to 192 (some of the ensuing homework problems are based on examples from this section).

Solution NA.

\vspace{0.1in}
\noindent 2. A typesetter makes, on average, one mistake per 1000 word.  Let $X$ be the number of mistakes that he makes on a page with $n$ words.
\begin{itemize}
\item[a.] Suppose we model $X$ with a binomial distribution. What are the model parameters? Discuss the binomial model assumptions in the context of the problem.
\item[b.] With a binomial model, calculate the probabilities that: (i) he makes no mistake on a page with $100$ words; (ii) he makes exactly 1 mistake on a page with $100$ words; (iii) he makes 1 or more mistake on a page with $100$ words; (iv) he makes no mistake on a page with $200$ words.
\item[c.] Suppose we model $X$ with a poisson distribution. What is the model parameter? Discuss the poisson model assumptions in the context of the problem.
\item[d.] Re-calculate the probabilities in b using a poisson model.
\item[e.] Simulate $1000$ random observations from the binomial model in a and plot a histogram of these observations; simulate another $1000$ random observations from the poisson model in c and also plot a histogram. Compare the binomial and poisson models by comparing the two histograms. 
\end{itemize}

Solution: We can model $X$ with binomial ($n$, $p=0.001$). We assume that (1) whether he makes a mistake on one word is \textit{independent} from whether he makes mistakes on other words, and (2) each word has \textit{equal probability} that he makes a mistake. Probabilities in b are calculated below.
```{r}
dbinom(0,size=100,prob=0.001)
dbinom(1,size=100,prob=0.001)
1-dbinom(0,size=100,prob=0.001)
dbinom(0,size=200,prob=0.001)
```

We can also model $X$ with poisson ($\lambda=0.001n$). This is an approximation because the sample space of number of mistakes he can make is $0, 1, \ldots, n$, but the poisson distribution has sample space $0,1,\ldots,\infty$. We assume that (1) whether he makes a mistake on one word is \textit{independent} from whether he makes mistakes on other words, and (2) he makes mistakes at a \textit{constant rate}.   Probabilities in d are calculated below. Note that the probabilities are very close to the binomial model.
```{r}
dpois(0,lambda=0.1)
dpois(1,lambda=0.1)
1-dpois(0,lambda=0.1)
dpois(0,lambda=0.2)
```

Histograms to compare the binomial and poisson models are given below. Comment: it's hard to estimate tail probabilities like $P(X=3,4,\ldots,100)$ using simulation because the probabilities are very small.  
```{r}
par(mfrow=c(1,2))
plot(as.factor(rbinom(1000,size=100,prob=0.001)),
     col='lightblue',main='binomial')
plot(as.factor(rpois(1000,lambda=0.1)),
     col='purple',main='poisson')
```


\vspace{0.1in}
\noindent 3. Do question 18  (page 199) in \textit{Introduction to Probability}.

Solution: 
\begin{itemize}
\item[(a)] The number of  raisins in a randomly picked cookie, $X$, can be modeled using a poisson distribution with $\lambda=600/500$, so $P(X=0)$ is \texttt{dpois(0,lambda=6/5)}=`r dpois(0,lambda=6/5)`.
\item[(b)] Similar to (a), we use poisson with $\lambda=400/500$.\\
$P(X=2)$ = \texttt{dpois(2,lambda=4/5)}=`r dpois(2,lambda=4/5)`.
\item[(c)]$P(X\ge2)=1-P(X=0,1)$. \\
\texttt{1-dpois(0,lambda=1000/500)-dpois(1,lambda=1000/500)}=`r 1-dpois(0,lambda=1000/500)-dpois(1,lambda=1000/500)`.
\end{itemize}

\vspace{0.1in}
\noindent 4. Under the setup of question 26 (page 201) in \textit{Introduction to Probability}, the number of hits in each of the $576$ small areas can be modeled by a poisson distribution. What is the parameter? Calculate the probabilities that a small area would have 0, 1, 2, 3, 4, and 5 or more hits under the poisson model, and compare the probabilities to the actual proportions.

Solution: The number of hit can be modeled by poisson with $\lambda=537/576$.  Calculation below shows that the poisson model works very well.
```{r}
#estimated probabilities
estimated=dpois(c(0,1,2,3,4),lambda=537/576) 
estimated[6]=1-sum(dpois(c(0,1,2,3,4),lambda=537/576))

#actual proportions
actual=c(229,211,93,35,7,1)/537

result=cbind(estimated,actual)
rownames(result)=c('0 hit','1 hit','2 hits','3 hits','4 hits', '5 or more hits')
round(result,3)
```

\vspace{0.1in}
\noindent 5. Do question 33  (page 202) in \textit{Introduction to Probability} by simulating the process and making random draws from the distribution of $X$. The random draws can be used to empirically study the distribution of $X$. For instance, plot a histogram of the random draws to estimate the shape of the distribution; use observed proportions to estimate the probability mass function.

Solution: Simulation given below. We can estimate, for instance, $P(X=0)=0.05,P(X=1)=0.15$,etc.
```{r}
nsim=10000
result=rep(NA,nsim)

for(i in 1:nsim){
  numcar=rpois(1,lambda = 4) #total number of cars
  result[i]=rbinom(1,size=numcar,prob=3/4) #number of cars that choose Main Street
}
table(result)/nsim #estimated probabilities
plot(as.factor(result),main='Number of cars that go to Main Street',col='salmon')
```

\vspace{0.1in}
\noindent 6. Suppose that the probabilities of having a male or a female child are both 0.5, and you do not have to consider cases like twins. Calculate the following probabilities using appropriate models.
\begin{itemize}
\item[a.] If you plan to have three children, what is the probability that you have exactly one daughter? What is the probability that you have at least one daughter?
\item[b.] What is the probability that your first daughter is your third child? (Hint: imagine the scenario of keeping having children until you have a daughter, and model the number of children when the first daughter ``occurs'').
\item[c.] What is the probability that your second daughter is your third child?
\item[d.] How many children do you need to have to guarantee with more than 95\% chance that you will have at least a daughter?
\end{itemize}

Solution: For a, b, c:
```{r}
dbinom(1,size=3,prob=0.5)  #a, exactly 1
1-dbinom(0,size=3,prob=0.5)  #a, at least 1

dgeom(2,prob=0.5) #b

dnbinom(1,size=2,prob=0.5) #c
```

For d, let N be the number of children you have, then the number of daughters, $X$, can be modeled as binomial with $n=N$ and $p=0.5$. We want $P(X\ge1)=1-P(X=0)$ to be greater than $95\%$. We calculate this probability for different values of $N$, and see that $N=5$ is the smallest number of children to satisfy the condition.
```{r}
1-dbinom(0,size=c(1:10),prob=0.5)
```






